---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"), header=TRUE ,sep=",")
unlink("acitivity.zip")

```
## What is mean total number of steps taken per day?
 
```{r}
daytotal<- tapply(data$steps,data$date,sum,na.rm=TRUE)
hist(daytotal, main = "Total Number of Steps taken per day", xlab= "", col="powderblue")
meansteps<-format(round(mean(daytotal),0),scientific = FALSE)
mediansteps<-format(round(median(daytotal),0),scientific = FALSE)
```

The mean total number of steps taken per day is `r meansteps`.  
The median total number of steps taken per day is `r mediansteps`.

## What is the average daily activity pattern?
```{r}
intervalmean<-tapply(data$steps,data$interval,mean,na.rm=TRUE)
intervals<-as.numeric(names(intervalmean))
plot(intervals,intervalmean,type="l",xlab="5-minute interval",ylab="average number of steps taken")
maxinterval<- names(intervalmean)[intervalmean ==max(intervalmean)]
```

On average across all the days in the dataset, interval `r maxinterval` contains the maximum number of steps.


## Imputing missing values
```{r}
missing<- sum(is.na(data$steps))
```
Total number of missing values in the dataset is `r missing`.

If a value is missing, I will use the mean for that 5-minute interval to fill in.
```{r}
obs<- nrow(data)
newsteps<-numeric(obs)

for (i in 1:obs){
  if(is.na(data$steps[i]))
    newsteps[i]=as.numeric(intervalmean[as.numeric(names(intervalmean))==data$interval[i]])
  else
    newsteps[i]=data$steps[i]
}

date<-data$date
interval<-data$interval
newdata<-data.frame(newsteps,date,interval)
newdaytotal<- tapply(newsteps,date,sum)
hist(newdaytotal, main = "New Total Number of Steps taken per day", xlab= "", col="rosybrown1")
newmean<-format(round(mean(newdaytotal),0),scientific = FALSE)
newmedian<-format(round(median(newdaytotal),0),scientific = FALSE)

```
The mean total number of steps taken of the new dataset is `r newmean`.  
The median total number of steps taken of the new dataset is `r newmedian`.   
Those values are different from the first part of the assignment. Imputing missing data increased the estimates of the total daily number of steps.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(dplyr)
newdata1<-mutate(newdata,group=ifelse(weekdays(as.Date(newdata$date))=="Sunday"|weekdays(as.Date(newdata$date))=="Saturday","weekend","weekday"))
newdata1$group<-as.factor(newdata1$group)
newdata2<-group_by(newdata1, group,interval)
wkdy<-summarize(newdata2,mean(newsteps))
names(wkdy)<-c("group","interval","Mean")

library(lattice)
xyplot(wkdy$Mean ~ wkdy$interval | group, data = wkdy, layout= c(1,2),type="l",xlab="Interval",ylab="Number of steps")

```